---
output: github_document

always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# c3plot



The goal of c3plot is provide to a base 'graphics'-like interface to the [C3.js](https://c3js.org/) JavaScript charting library. The main `c3plot()` function is an S3 generic like the base `plot()` function with methods for various base R objects.

Besides `htmlwidgets`, it has minimal R dependencies. Tidyverse not required!

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("arkraieski/c3plot")
```
## Example

Here’s some example code for a basic scatter plot using R’s `airquality` dataset:

```{r examplePlot}
library(c3plot)
air <- airquality
c3plot(air$Ozone, air$Solar.R, 
       main = "New York Air Quality Measurements", 
       xlab = "Ozone (ppb)",
       ylab = "Solar Radiation (lang)",
       col = "red")

```


## Methods

To see a list of S3 methods:

```{r}
library(c3plot)
methods(c3plot)
```

More are on the way! The goal is to replicate the behavior of base R `plot()` methods as closely as possible while also getting the benefits of using C3 (ie. interactivity). Each method has its own help page that can be accessed for additional information.


## Testing

Basic tests are implemented, but this is a WIP. The package is tested on Windows, Linux, and MacOS.
